Network Configuration Synthesis with Abstract Topologies

### 2016-11-10 paper/abstraction/paper.tex
add citations for topologies + Z3

\newcommand{\inferencetwo}[1]{
    \node[draw, anchor=west] at (#1 + 2, 1.25) {$e_i > 0$};
    \node[anchor=east] at (#1 - .4, 1.25) {\textbf{I-striping}};
    \node at (#1 - 1 - .11, .6) {$e_1$};
    \node at (#1 + .11, 1.5) {$e_2$};
    \node at (#1 + 2 + .11, .6) {$e_3$};
    \node at (#1 + .11 + .8, 1.5) {$e_4$};
    \node at (#1 - 1, -.8) {L$(j,k)$};
    \node at (#1 +. 5, 2.9) {S$(min(j*k,g),1)$};
    \node at (#1 + 2, -.8) {S$\big( min(j, g *\frac{e_4}{e_3}),1\big)$};
    \node at (#1, 3.3) {};
    \draw [] (#1 + .5, 2.01) circle [radius=0.4] node {$n$};
    \draw [] (#1 - 1, 0) circle [radius=0.4] node {$m$};
    \draw [] (#1 + 2, 0) circle [radius=0.4] node {$o$};
    \draw [->] (#1 - 1, .4) -- (#1 + .5, 2 - .4);
    \draw [<-] (#1 + 2, .4) -- (#1 + .5, 2 - .4);
}

@@

\begin{figure*}[t!]
  \begin{minipage}[t]{\linewidth}%
  \begin{tikzpicture}
    \inference{0}{L$(j,k)$}{$e_1 > 0$}{S$(min(j*k,e_1),1)$}{I-out1}{$e_1$}{$e_2$}
    \inference{4.5}{A$(j,k)$}{$e_2 > 0$}{A$(1,min(j,e2))$}{I-out2}{$e_1$}{$e_2$}
    \inference{9}{L$(j,k)$}{$e_1=n$}{A$(min(j*k,n),1)$}{I-mesh1}{$e_1$}{$e_2$}
    \inference{13.5}{L$(j,k)$}{$e_1=n$}{A$(1,j)$}{I-mesh2}{$e_1$}{$e_2$}
  \end{tikzpicture}
  \end{minipage}

@@

The abstract disjoint path analysis starts from a fixed source location $src$ and repeatedly applies the inference rules from Figure~\ref{fig:inference-rules} until it reaches a fixed point. Inference rule conditions are checked for validity over all concrete topologies using the Z3 \ryan{cite} SMT solver. In order to ensure termination, we make the observation that, for any fact learned of the form $L(j,k)$, it is sound to instead infer $L(j',k')$ if $j' \leq j$ and $k' \leq k$. Therefore, for each inference rule that learns a fact of the form $L(j,k)$ we minimize the symbolic expressions for $j$ and $k$ subject to the topology constraints using the optimization framework Z3Opt.

@@

Consider again the datacenter example from \S\ref{sec:overview}. Suppose now we add the valleyfree constraint that traffic never takes a path with a down-up in it for the routing policy for global prefixes:
%
\begin{code}
\$GP  \Path \End(TG) \ensuremath{\cap} \Novalley(\{TG,TL\},\{AG,AL\},\{S\})
\end{code}
\noindent
%
Figure~\ref{fig:example-inference} shows the abstract product graph representation for this routing policy. The inference algorithm starts from the node $(TG,0,0)$ with the initial fact $S_P S_{TG}(1,\infty)$ (\IE, no restriction on the number of disjoint paths initially).
%
The first step applies the rule \infrule{I-mesh1} together with \infrule{I-local} to reason about connectivity within a single pod for the TG and AG roles. This invokes a call to Z3Opt to minimize the expression $min(\infty, AG)$, which results in 2. Therefore, the algorithm learns a new invariant of the form $S_P A_{AG} (2,1)$ for node $(AG, 0)$ to indicate that in some pod P, any group of 2 nodes is reachable.
%
The next step uses the inference rule \infrule{I-out2} to learn that any single spine node is reachable. It also can apply rule \infrule{I-striping} to determine that there is some group of at least 2 spine routers reachable at node $(S,0)$ and also that there is some group of at least 2 nodes reachable in the AL role in state $(AL,0)$.

@@

The \sysname compiler is implemented in approximately 9500 lines of F\# code. The compiler accepts both concrete and abstract topologies and will generate router configurations for open-source Quagga routers. Operators can specify their required fault tolerance level for aggregation safety. The abstract safety analysis uses Z3Opt \ryan{reference} to minimize variables subject to the annotated constraints. Since the analysis typically calls Z3 with many times with relatively simple optimization problems, we use a timeout of 200ms.

@@

\begin{figure}[t!]
  \begin{center}
      \begin{tabular}{| l | c | c | c |}
      \hline
      \textbf{Topology} & \textbf{Fixed} & \textbf{Reachable} & \textbf{K-paths} \\ \hline
      Google Fattree & & \cmark & \cmark  \\ \hline
      Facebook Fattree & & \cmark & \cmark \\ \hline
      Microsoft Fattree & & & \\ \hline
      F10 Fattree & & \cmark & \cmark \\ \hline
      BCube & $k$ & \cmark & \xmark \\ \hline
      DCell & $k$ & \cmark & \xmark \\ \hline
         %& HCN & $h$ & \cmark & \xmark \\
      Butterfly & $n$ & \cmark & \cmark \\ \hline
      Hypercube & $N$ & \cmark & \cmark \\ \hline
      HyperX & $L$ & \cmark & \cmark \\ \hline
      \end{tabular}
  \end{center}
  \caption{Abstract Analysis Precision.}
  \label{fig:analysis-precision}
\end{figure}

@@

The Fat Tree topology is one of the most widely used datacenter topologies. We look at the precision of our analysis on three variants of the Fat Tree~\cite{foo} topology used in practice: the Google Fat Tree, the Facebook Fat Tree, and the F10 fault tolerant Fat Tree. For each Fat Tree variant, we use a tiered abstraction similar to that in our previous example \ryan{Reference}. Each Fat Tree topology is parameterized over the number of pods $k$, which can be scaled up in the abstraction to allow for expansion. For all Fat Tree variants, the analysis is precise enough to determine the exact number of disjoint paths between all pairs of concrete routers.

@@

We also model several recursive topologies using our abstractions. These include the BCube and DCell topologies. Each topology includes a recursion depth parameter ($k$), which we must fix. For a recursive topology with depth $d$, we model it as an abstract topology consisting of abstract nodes for each of the depth $d-1$ subcomponents. This allows for safe expansion within a subcomponent, but not from increasing the recursion depth. The analysis is able to accurately determine reachability (i.e., at least 1 disjoint path) for each recursive topology, but is overly conservative in determining k-disjoint paths. For the BCube topology, it is able to to determine the correct number of disjoint paths between some pairs of nodes, but not all.

@@

Hypercube variants can be used as an alternative to Clos-style topologies for networks with high-radix (port density) routers. The HyperX topology can be viewed as a generalization of the hypercube, which includes parameters $L$ for the lattice dimension of the network, $S_i$ for the node multiplicity of each dimension $i$, $K_i$ for the bandwidth of links for each dimension, and $T$ for the number of terminal nodes (servers) connected to each router. For a fixed number of dimensions $L$, we abstract each full mesh of $S_L$ nodes into its own abstract node. The $S_{x-1}$ groups of abstract nodes can be captured using pods of the abstract $S_x$ nodes.

>>>>

\newcommand{\inferencetwo}[1]{
    \node[draw, anchor=west] at (#1 + 2, 1.25) {$e_i > 0$};
    \node[anchor=east] at (#1 - .4, 1.25) {\textbf{I-striping}};
    \node at (#1 - 1 - .11, .6) {$e_1$};
    \node at (#1 + .11, 1.5) {$e_2$};
    \node at (#1 + 2 + .11, .6) {$e_3$};
    \node at (#1 + .11 + .8, 1.5) {$e_4$};
    \node at (#1 - 1, -.8) {L$_m(j,k)$};
    \node at (#1 +. 5, 2.9) {S$_n(min(j*k,g),1)$};
    \node at (#1 + 2, -.8) {S$_o\big( min(j, g *\frac{e_4}{e_3}),1\big)$};
    \node at (#1, 3.3) {};
    \draw [] (#1 + .5, 2.01) circle [radius=0.4] node {$n$};
    \draw [] (#1 - 1, 0) circle [radius=0.4] node {$m$};
    \draw [] (#1 + 2, 0) circle [radius=0.4] node {$o$};
    \draw [->] (#1 - 1, .4) -- (#1 + .5, 2 - .4);
    \draw [<-] (#1 + 2, .4) -- (#1 + .5, 2 - .4);
}

@@

\begin{figure*}[t!]
  \begin{minipage}[t]{\linewidth}%
  \begin{tikzpicture}
    \inference{0}{L$_m(j,k)$}{$e_1 > 0$}{S$_n(min(j*k,e_1),1)$}{I-out1}{$e_1$}{$e_2$}
    \inference{4.5}{A$_m(j,k)$}{$e_2 > 0$}{A$_n(1,min(j,e2))$}{I-out2}{$e_1$}{$e_2$}
    \inference{9}{L$_m(j,k)$}{$e_1=n$}{A$_n(min(j*k,n),1)$}{I-mesh1}{$e_1$}{$e_2$}
    \inference{13.5}{L$_m(j,k)$}{$e_1=n$}{A$_n(1,j)$}{I-mesh2}{$e_1$}{$e_2$}
  \end{tikzpicture}
  \end{minipage}

@@

The abstract disjoint path analysis starts from a fixed source location $src$ and repeatedly applies the inference rules from Figure~\ref{fig:inference-rules} until it reaches a fixed point. Inference rule conditions can be checked for validity over all concrete topologies using an SMT solver. In order to ensure termination, we make the observation that, for any fact learned of the form $L(j,k)$, it is sound to instead infer $L(j',k')$ if $j' \leq j$ and $k' \leq k$. Therefore, for each inference rule that learns a fact of the form $L(j,k)$ we minimize the symbolic expressions for $j$ and $k$ subject to the topology constraints using the optimization framework $\nu$Z~\cite{z3opt}.

@@

Consider again the datacenter example from \S\ref{sec:overview}. Suppose now we add the valleyfree constraint that traffic never takes a path with a down-up in it for the routing policy for global prefixes:
%
\begin{code}
\$GP  \Path \End(TG) \ensuremath{\cap} \Novalley(\{TG,TL\},\{AG,AL\},\{S\})
\end{code}
\noindent
%
Figure~\ref{fig:example-inference} shows the abstract product graph representation for this routing policy. The inference algorithm starts from the node $(TG,0,0)$ with the initial fact $S_P S_{TG}(1,\infty)$ (\IE, no restriction on the number of disjoint paths initially).
%
The first step applies the rule \infrule{I-mesh1} together with \infrule{I-local} to reason about connectivity within a single pod for the TG and AG roles. This invokes a call to $\nu$Z to minimize the expression $min(\infty, AG)$, which results in 2. Therefore, the algorithm learns a new invariant of the form $S_P A_{AG} (2,1)$ for node $(AG, 0)$ to indicate that in some pod P, any group of 2 nodes is reachable.
%
The next step uses the inference rule \infrule{I-out2} to learn that any single spine node is reachable. It also can apply rule \infrule{I-striping} to determine that there is some group of at least 2 spine routers reachable at node $(S,0)$ and also that there is some group of at least 2 nodes reachable in the AL role in state $(AL,0)$.

@@

The \sysname compiler is implemented in approximately 9500 lines of F\# code. The compiler accepts both concrete and abstract topologies and will generate router configurations for open-source Quagga routers. Operators can specify their required fault tolerance level for aggregation safety. The abstract safety analysis uses $\nu$Z~\cite{z3opt} test validity and to minimize variables subject to the annotated constraints. Since the analysis typically calls the SMT solver with many times with relatively simple optimization problems, we use a timeout of 200ms.

@@

\begin{figure}[t!]
  \begin{center}
      \begin{tabular}{| l | c | c | c |}
      \hline
      \textbf{Topology} & \textbf{Fixed} & \textbf{Reachable} & \textbf{K-paths} \\ \hline
      Fattree~\cite{fattree} & & \cmark & \cmark  \\ \hline
      Facebook Fattree~\cite{facebook-fattree} & & \cmark & \cmark \\ \hline
      F10 Fattree~\cite{f10-fattree} & & \cmark & \cmark \\ \hline
      VL2 Fattree~\cite{vl2-fattree} & & & \\ \hline
      BCube~\cite{bcube} & $k$ & \cmark & \xmark \\ \hline
      DCell~\cite{dcell} & $k$ & \cmark & \xmark \\ \hline
         %& HCN & $h$ & \cmark & \xmark \\
      Butterfly~\cite{butterfly} & $n$ & \cmark & \cmark \\ \hline
      Hypercube & $N$ & \cmark & \cmark \\ \hline
      HyperX~\cite{hyperx} & $L$ & \cmark & \cmark \\ \hline
      \end{tabular}
  \end{center}
  \caption{Abstract Analysis Precision.}
  \label{fig:analysis-precision}
\end{figure}

@@

The Fat Tree topology is one of the most widely used datacenter topologies. We look at the precision of our analysis on four variants of the Fat Tree topology used in practice: a Standard Fat Tree~\cite{fattree}, the Facebook Fat Tree~\cite{facebook-fattree}, the F10 fault tolerant Fat Tree~\cite{f10-fattree}, and the VL2 topology~\cite{vl2-fattree}. For each Fat Tree variant, we use a tiered abstraction similar to that in our previous example. Each Fat Tree topology is parameterized over the number of pods $k$, which can be scaled up in the abstraction to allow for expansion. For all Fat Tree variants, the analysis is precise enough to determine the exact number of disjoint paths between all pairs of concrete routers.

@@

We also model several recursive topologies using our abstractions. These include the BCube~\cite{bcube} and Dcell~\cite{dcell} topologies. Each topology includes a recursion depth parameter ($k$), which we must fix. For a recursive topology with depth $d$, we model it as an abstract topology consisting of abstract nodes for each of the depth $d-1$ subcomponents. This allows for safe expansion within a subcomponent, but not from increasing the recursion depth. The analysis is able to accurately determine reachability (i.e., at least 1 disjoint path) for each recursive topology, but is overly conservative in determining k-disjoint paths. For the BCube topology, it is able to to determine the correct number of disjoint paths between some pairs of nodes, but not all.

@@

Hypercube variants can be used as an alternative to Clos-style topologies for networks with high-radix (port density) routers. The HyperX~\cite{hyperx} topology can be viewed as a generalization of the hypercube and butterfly topologies, which includes parameters $L$ for the lattice dimension of the network, $S_i$ for the node multiplicity of each dimension $i$, $K_i$ for the bandwidth of links for each dimension, and $T$ for the number of terminal nodes (servers) connected to each router. For a fixed number of dimensions $L$, we abstract each full mesh of $S_L$ nodes into its own abstract node. The $S_{x-1}$ groups of abstract nodes can be captured using pods of the abstract $S_x$ nodes.
